---
title: "Introduction plpdf package"
author: "Aris Lourens"
output:
   rmarkdown::html_vignette:
      toc: true
      number_sections: false
vignette: >
  %\VignetteIndexEntry{Introduction plpdf package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introdcuction

This package makes calculations with random variables feasible. 
Random variables can be characterized by a probability distribution. 
In this package, continues random variables are approximated 
by piecewise-linear probability density functions (`PL-PDFs`).

Applications and details about the development of this package are found [here][thesis].

## Predefined distributions

Functions are available to define PL-PDFs of standard distributions:
uniform distribution `plUnif`,
trapezoid shaped distributions `plTrap`,
normal or Gaussian distribution `plNorm`, and
the log-normal distributions `plLnorm`.



## Options

The `PL-PDFs` are discretized in a number of bins with a default value of 50.
Usually, this gives a reasonable accurate result.
If more accuracy is needed this number can be increased using [plpdfSetNbin](../help/plpdfSetNbin).
If faster calculations are needed this number can be decreased.


## Class plpdf

The probability density functions are defined as objects of class `plpdf`.
It is defined as a list containing an `x`-array, which are the domain values 
of the function, 
and a `y`-array, which defines the probability densities.
An optional array `Y` may be available containing the cumulative probability.

## S3 methods for generic functions

### plot

The [plot()] method is available for `plpdf` objects.
Options useful for PDFs are available, 
like limiting the domain (x-axis) for heavy tailed PDFs 
or drawing the cumulative distribution.


### Group generic functions

For some of the [Group Generic functions](../../base/help/groupGeneric) methods 
are defined for class `plpdf` objects.
Not all functions as defined in R make sense when applied to probability density functions.
For now, the Complex group is omitted as a whole.

#### Math

Methods for the functions `log`, `exp` and `sqrt` are available.


#### Ops

This group defines binary operations. 
Methods are supplied for the operators `+, -, *, /, ^`.


#### Summary

The implemented generic functions are `sum`, `min`, `max`, `range`.

`sum` returns the integral (probaility) of the PL-PDF function. If multiple pdf's are 
given then the total sum is returned. 
Usually, this should be 1 for a single PL-PDF but this can diverge from 1 caused by 
roundoff errors.

`min`, `max` and `range` return the respective values of the domain (`x`-value) of the pdf's.


### Other functions

The function `mean` returns the mean of the PL-PDF.

Unfortunately, the function [var](../../stats/help/var) is not a generic function within R. 
Therefore, the function [plVar](../help/plVar) is added. 


The function `is.na` checks the density array of a `plpdf` object for `NA`-values. 
In a valid `plpdf` object this should never be the case.

The function [as.plpdf](../help/as.plpdf) transforms a list with `x` and `y` arrays into a PL-PDF.

```{r setup}
library(plPDF)
```

[thesis]: https://doi.org/10.33540/698

